---
title: "Stat 156 Summary Stats/Table A1"
format: pdf
editor: visual
---

```{r}
# Load required libraries
library(dplyr)
library(broom)

# Load the data
data <- haven::read_dta("table_A1.dta")

# Define the sample indicator and numeric variables
sample_col <- "sample" # Assuming this is the binary column for In/Not In Sample
numeric_vars <- names(data)[sapply(data, is.numeric) & names(data) != sample_col]

# Function to calculate summary statistics and t-tests and includes constant variables- in order for the difference in means t-test to make sense
calculate_summary <- function(data, variable, sample_col) {
  group_stats <- data %>%
    group_by(!!sym(sample_col)) %>%
    summarize(
      mean = mean(!!sym(variable), na.rm = TRUE),
      se = sd(!!sym(variable), na.rm = TRUE) / sqrt(n()),
      sd = sd(!!sym(variable), na.rm = TRUE),
      median = median(!!sym(variable), na.rm = TRUE),
      iqr = IQR(!!sym(variable), na.rm = TRUE)
    )
  
  # Check if standard deviation is non-zero for both groups
  if (any(group_stats$sd == 0 | is.na(group_stats$sd))) {
    return(tibble(
      Variable = variable,
      `Mean (In Sample)` = NA,
      `SE (In Sample)` = NA,
      `Median (In Sample)` = NA,
      `IQR (In Sample)` = NA,
      `Mean (Not In Sample)` = NA,
      `SE (Not In Sample)` = NA,
      `Median (Not In Sample)` = NA,
      `IQR (Not In Sample)` = NA,
      `Difference in Means` = NA,
      `P-value` = NA
    ))
  }
# Calculate t-test  
  t_test <- t.test(
    data[[variable]][data[[sample_col]] == 1],
    data[[variable]][data[[sample_col]] == 0]
  )
 #Output desired table  
  tibble(
    Variable = variable,
    `Mean (In Sample)` = group_stats$mean[group_stats[[sample_col]] == 1],
    `SE (In Sample)` = group_stats$se[group_stats[[sample_col]] == 1],
    `Median (In Sample)` = group_stats$median[group_stats[[sample_col]] == 1],
    `IQR (In Sample)` = group_stats$iqr[group_stats[[sample_col]] == 1],
    `Mean (Not In Sample)` = group_stats$mean[group_stats[[sample_col]] == 0],
    `SE (Not In Sample)` = group_stats$se[group_stats[[sample_col]] == 0],
    `Median (Not In Sample)` = group_stats$median[group_stats[[sample_col]] == 0],
    `IQR (Not In Sample)` = group_stats$iqr[group_stats[[sample_col]] == 0],
    `Difference in Means` = t_test$estimate[1] - t_test$estimate[2],
    `P-value` = t_test$p.value
  )
}

# Re-run for all numeric variables
summary_table <- purrr::map_dfr(numeric_vars, ~calculate_summary(data, .x, sample_col))

# Print the summary table
print(summary_table)

# Optionally save to CSV
write.csv(summary_table, "summary_table.csv", row.names = FALSE)

```

We replicate Table A (Mean Characteristics of Cities In and Out of Sample) from the paper.

While our table isn't exactly identical to Table A from the paper, but the values- both means and p-values are very similar (at least to the 1st decimal place).

The covariate means, overall look very balanced between the sampled and out-of-sample observations. For the few variables (percent black-1940 and Concentration index–1990) where the p-value is significant (at 5% level), the overall difference in means in our table are close to 0 (difference in mean of 0.1 for the concentration index, and -0.018 for percentage black 1940). Thus, overall, it seems that the observations/cities are pretty well-balanced between sample and not-in-sample observations.

```{r}
# Customize options
library(xtable)
latex_table <- xtable(data)
print(latex_table, type = "latex", include.rownames = TRUE)

```
